---
title: "Challenge 7: Incorporating Multiple Inputs"
subtitle: "Functions + Fish"
author: "Gabriela Tena!"
format: 
  html:  
    embed-resources: true
    code-tools: true
    toc: true
    code-fold: true
editor: source
execute: 
  echo: true
warning: false
message: false
theme: "solar"
---

```{r}
#| label: setup
library(tidyverse)

fish <- read_csv('https://raw.githubusercontent.com/gabytena/lab-7/refs/heads/main/data/BlackfootFish.csv')

```

## Condition Index

A frequently used measurement for fish health is a condition index [(Wikipedia article)](https://en.wikipedia.org/wiki/Condition_index_in_fish). The following
simple equation can be used to calculate the approximate condition index of a
fish:

$$\text{condition index} = \frac{weight}{length^3} \times 100$$

**1. There are specific units required for the calculation of a condition index -- length must be in centimeters and weight must be in grams. The weight data for the Blackfoot River fish were collected in grams; the length data were collected in millimeters. Transform the length data to the correct units.**

```{r}
#| label: transform-data-to-have-correct-units
fish <- fish |>
  mutate(length = length / 10)
```

**2. Collecting data of this sort can be very messy! Write a function that will replace unlikely `length` and `weight` measurements with `NA`. Your function should accept at least three inputs:**

+ **a vector of measurements,**
+ **the minimum reasonable value,**
+ **the maximum reasonable value.**

**If a value falls outside these bounds, you should replace it with an `NA`.**

::: callout-tip
If you are struggling with the structure of your function, I would suggest reading the [Mutating Function](https://r4ds.hadley.nz/functions.html#mutate-functions) from R4DS.
:::

```{r}
#| label: function-to-change-impossible-measurements-to-NA
unlikely_vals <- function(vec, min, max){
  likely_vals <- ifelse(vec < min | vec > max, 
                        NA,
                        vec)
    
}
```

**3. After consulting the [Montana Record Table](https://fwp.mt.gov/fish/anglingData/records) for the four species of trout included in these data, I have conjectured that it is unlikely to have measurements for fish below 5 cm and above 80 cm in length or below 10 g and above 4,000 g in weight. Use your function to modify the `length` and `weight` columns of the `fish` dataset based on my cutoffs.**

```{r}
#| label: modify-impossible-values-of-length-weight
fish |>
  mutate(
    length = unlikely_vals(length,5, 80),
    weight = unlikely_vals(weight, 10, 4000)
  )
```

**3. Write a function to calculate the condition index of a fish, given inputs of weight and length.**

::: callout-warning
Consider whether you will write a function that works with vectors (accepting vectors as inputs) or a function that works with data frames (accepting a dataset and variable names as inputs)!
:::

```{r}
#| label: condition-function
condition_idx <- function(df, c_weight, c_len) {
  df <- df |>
    mutate(condition_index = (df[[c_weight]] / (df[[c_len]]^3)) * 100)
}
```

**4. Make a thoughtful visualization of how fish conditions have varied over the duration of this study.**

::: callout-warning
### To be thoughtful...

... you need to use at least a third variable!
:::

```{r}
#| label: condition-indices-over-time
fish |>
  condition_idx( "weight", "length") |>
  ggplot(aes(x = as.factor(year),
             y = condition_index, 
             fill = species)) +
  geom_boxplot(outlier.shape = NA) +
  labs(
    title = "Condition Index Distribution Over Time by Species",
    x = "Year",
    y = "",
    fill = "Species"
  ) +
  scale_y_continuous(limits = c(.5,
                                1))+
  theme(legend.position = "top")+
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")

```
